Reasoning About a Machine with Local Capabilities

### 2019-01-25 toplas/paper.tex
Address reviewer 3 concerns
Address the remaining of TOPLAS reviewer comments, first pass

\definecolor{OliveGreen}{RGB}{60,128,49}
\definecolor{Blue}{RGB}{92,26,143}
\newenvironment{toplas}{}{}
    {\color{OliveGreen}}{}
\newenvironment{toplassug}
    {\begin{comment}}{\end{comment}}

\newcommand{\itoplas}[1]{#1}
%    {{\color{OliveGreen} #1}}
\newcommand{\itoplassug}[1]{}
%    {{\color{Blue} #1}}

@@

A similar problem is the enforcement of control-flow correctness on capability
machines. An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}:
the standard contiguous C stack is split into a central, trusted stack, managed
by trusted call and return instructions, and disjoint, private, per-compartment
stacks. To prevent illegal use of stack references, the approach relies on
\emph{local capabilities}, a type of capabilities offered by CHERI to
\emph{temporarily} relinquish authority, namely for the duration of a function
invocation whereafter the capability can be revoked. However, details are scarce
(how does it work precisely? what features are supported?) and a lot remains to
be investigated (e.g., combining disjoint stacks with cross-domain function
pointers seems like it will scale poorly to large numbers of components?).
Finally, there is no argument that the approach is watertight and it is not even
clear what security property is targeted exactly.

@@

\begin{figure}[hb]
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl l rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
     && \Caps  &\isdef & \multicolumn{6}{l}{\left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}}
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lt{\lv}{\rv}{\rv} \mid
                \lea{\lv}{\rv} \mid 
                \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid \\
          &   &  \getp{\lv}{\lv} \mid 
                 \getb{\lv}{\lv} \mid
                 \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

\begin{figure}[htb]
  \centering
  \begin{equation*}
    \Phi  \rightarrow
    \begin{cases}
      \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
      \begin{array}[t]{l}
        \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ 
        \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
      \end{array}
\\
      \failed                                 & \text{otherwise}
    \end{cases}
  \end{equation*}
  \begin{align*}
    \stdUpdatePc{\Phi} &=
                         \begin{cases}
                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
                           \begin{array}[t]{l}
                             \text{if $\memreg(\pcreg) = \stdcap$}\\
                             \text{\;\;and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
                           \end{array}
\\
                             \failed & \text{otherwise}
                         \end{cases}
  \end{align*}
  \begin{tabular}{|c|p{3.4cm}|>{\raggedright\arraybackslash}p{7.3cm}|}
    \hline
    $i$&$\sem{i}(\Phi)$&Conditions\\
    \hline 
    $\fail$&$\failed$&\\
    \hline
    $\halt$&$(\halted,\memheap)$&\\
    \hline
    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$r_2 \in \Regs \Rightarrow w = \memreg(r_2)$ and $r_2 \in \ints \Rightarrow w = r_2$\\
    \hline
    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and  $w = \memheap(\addr)$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
    \hline
    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
    \hline
    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
    \hline
    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
                                                                              and $(\perm',g') \sqsubseteq (\perm,g)$ and  $w =((\perm',g'),\start,\addrend,\addr)$\\
    \hline
    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_1,n_2,\addr)$}\\
    \hline
    \itoplas{$\lea{\refreg{r_1}}{r_2}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{c}}}$} &\itoplas{$\memreg(r_1) = \stdcap$ and $n = \memreg(r_2)$ and $n \in \ints $ and $\perm \neq \entry$ and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$} \\
    \hline
    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
    \hline
    \multicolumn{3}{|c|}{$\cdots$}\\
    \hline
    \_&$\failed$&otherwise\\
    \hline
  \end{tabular}
  \caption{An excerpt from the operational semantics.}
  \label{fig:op-sem}
\end{figure}

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
We assume an infinite set of addresses $\Addrs$ and define machine words as
either integers or capabilities of the form
$((\perm,\gl),\var{base},\var{end},\addr)$. Such a capability represents the
authority to execute permissions $\perm$ on the memory range
$[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag
$\gl$ indicating whether the capability is global or local. There is no notion
of pointers other than capabilities, so we will use the terms interchangeably.
The available permissions are null permission
($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute
($\exec$) and read/write/execute ($\rwx$) permissions. Additionally, there are
three special permissions: read/write-local ($\readwritel$), read/write-local/execute
($\rwlx$) and enter ($\entry$), which we will explain below.
\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}. We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}

@@

@@

\dominique{Follow order of figure in discussion?}
The $\load{}{}$ instruction loads the contents of the requested memory location
into a register, but only if the capability has appropriate authority (i.e.\
read permission and an appropriate range). $\restricttwo{}{}$ updates a
capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original. 
%\itoplassug{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
%\itoplassug{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
% Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
It also never turns local
capabilities into global ones. $\geta{}{}$ queries the current address of a
capability and stores it in a register.

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$ that is then in $\pcreg$. 
\lau{As we talked about what is described here, is not how we make closures in
the TR, so we should consider whether we should refrain from using that word
here. }

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we take care to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

\textbf{Program start-up} We assume that the language runtime initializes the
memory as follows: a contiguous array of memory is reserved for the stack, for
which we receive a stack pointer in a special register $r_\stk$. We stress that
the stack is not built-in, but merely an abstraction we put on this piece of the
memory. The stack pointer is local and has $\rwlx$ permission. Note that this
means that we will be placing and executing instructions on the stack.
Crucially, the stack is the only part of memory for which the runtime (including
malloc, loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
Additionally, our examples typically also assume some memory to store
instructions or static data. Another part of memory (called the heap) is
initially governed by malloc and at program start-up, no other code has
capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
assume that memory allocated through malloc cannot be freed.

@@

In one of the registers, we also provide a return pointer, which must be a local
capability. If it were global, the adversary would be able to store away the
return pointer in a global data structure (i.e.\ there exists a global capability for it), and jump to it later, in
circumstances where this should not be possible. For example, they could store
the return pointer, legally jump to it a first time, wait to be invoked again
and then jump to the old return pointer a second time, instead of the new return
pointer received for the second invocation. Similarly, they could store the
return pointer, invoke a function in our code, wait for us to invoke them again
and then jump to the old return pointer rather than the new one, received for the
second invocation.
% \lau{If we need space, then maybe one of the examples is sufficient.}
% \lau{Both of these are good to have their as the two last examples of \sectionname~\ref{sec:examples} use them.}
%
% \lau{In the examples, we refer to the local return pointer as
%   ``protected return pointer''. We should either introduce that here
%   or use a different formulation in the examples. }
% \lau{It seemed did not seem to add anything to adopt the terminlogy - we
%   often highlight that the return pointer is local which in most cases
%   is so important that we do not want to leave it out.}
By making the return pointer local, we prevent such attacks: the adversary can
only store local capabilities through write-local capabilities, which means (because
of our assumptions above): on the stack. Since the stack pointer itself is
also local, it can also only be stored on the stack. Because we clear the part
of the stack that the adversary has had access to before we pass control back,
there is no way for them to recover either of these local capabilities.

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores their stack pointer (\ref{fig:illu-adv-store-sp}), invokes trusted code
that uses part of the stack to store private data (\ref{fig:illu-adv-call}) and then invokes the adversary
again with a stack pointer restricted to exclude the part containing the private
data (\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
could access the private data stored there by the trusted code and break
local-state encapsulation.  
% \lau{Do we say anything about arguments to the callback? Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary. Can we even to any extend allow local arguments that we pass on from an unknown source?}
% Domi: this is now discussed in the discussion section, no need to discuss it here?

@@

Clearly, neither of these capabilities should be accessible by the adversary. In
other words, the return pointer provided to the adversary must be a capability
that they can jump to but not read from, i.e.\ an enter capability. To make this
work, we construct the activation record as depicted in
\figurename~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
over the entire activation record (containing the previous return and stack
pointer), and its current address points to a number of restore instructions in
the record, so that upon invocation, these instructions are executed and can
load the old stack pointer and program counter back into the register file. As
the return pointer is an enter pointer, the adversary cannot get hold of the
activation record's contents, but after invocation, its permission is updated to
$\exec$, so the contents become available to the restore instructions.

@@

\paragraph{Local capabilities}
A final thing we provide informal intuition for is how our results take into account local
capabilities and their special treatment by the hardware. Normally, when we
invoke an untrusted piece of code and provide it with certain global
capabilities, it may have stored those capabilities in memory and we will only
be able to reinvoke the code if we can guarantee that those values are still
valid. Formally, worlds represent the invariants that global capabilities'
safety relies on and the reinvocation is only safe in future worlds, where the
invariants are respected.

@@

      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
            {\fill[black] circle(1.5pt);}}}}
      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(g,l)$}} (5,7.25);

>>>>

\definecolor{OliveGreen}{RGB}{60,128,49}
\definecolor{Blue}{RGB}{92,26,143}
\newenvironment{toplas}%{}{}
    {\color{OliveGreen}}{}
\newenvironment{toplassug}
    {\begin{comment}}{\end{comment}}

\newcommand{\itoplas}[1]%{#1}
    {{\color{OliveGreen} #1}}
\newcommand{\itoplassug}[1]{}
%    {{\color{Blue} #1}}

@@

A similar problem is the enforcement of control-flow correctness on capability machines.
An interesting approach is taken in CheriBSD~\citep{Watson2015Cheri}: the standard contiguous C stack is split into a central, trusted stack and disjoint, private, per-compartment stacks.
The trusted stack is managed by trusted call and return instructions.
To prevent illegal use of stack references, the approach relies on \emph{local capabilities}, a type of capabilities offered by CHERI to \emph{temporarily} relinquish authority, namely for the duration of a function invocation whereafter the capability can be revoked.
However, details are scarce (how does it work precisely?
what features are supported?)
and a lot remains to be investigated (e.g., combining disjoint stacks with cross-domain function pointers seems like it will scale poorly to large numbers of components).
Finally, there is no argument that the approach is watertight and it is not even clear what security property is targeted exactly.

@@

We define the syntax of our capability machine in \figurename~\ref{fig:syntax}.
We assume an infinite set of addresses $\Addrs$ and define machine words as either integers or capabilities of the form $((\perm,\gl),\var{base},\var{end},\addr)$.
Such a capability represents the authority to execute permissions $\perm$ on the memory range $[\var{base},\var{end}]$, together with a current address $\addr$ and a locality tag $\gl$ indicating whether the capability is global or local.
There is no notion of pointers other than capabilities, so we will use the terms interchangeably.
The available permissions are null permission ($\noperm$), readonly ($\readonly$), read/write ($\readwrite$), read/execute ($\exec$) and read/write/execute ($\rwx$) permissions.
Additionally, there are three special permissions: read/write-local ($\readwritel$), read/write-local/execute ($\rwlx$) and enter ($\entry$), which we will explain below.
\itoplas{Permissions and locality are ordered and both orderings are depicted in \figurename~\ref{fig:perm-hier}.
  The write permission is subsumed by the write-local permission, i.e.\ a write-local capability can be used to store both $\local$ and $\glob$ capabilities to memory.
  We denote the pairwise ordering of permission and locality with $\sqsubseteq$.}

@@

\begin{figure}[b]
\[\arraycolsep=1.4pt
  \begin{array}{rlrcl l rlrcl}
    \addr   &\in& \Addrs &\isdef& \nats &\phantom{mak}                                & r    &\in & \RegName     &\bnfdef& \pcreg\mid r_0\mid r_1\mid\ldots\\
    w &\in&\Words &\isdef& \ints + \Caps &                                                   & \reg &\in & \Regs        &\isdef& \RegName \rightarrow \Words \\
    \perm   &\in& \Perms &::= & \noperm \mid \readonly\mid \readwrite\mid \readwritel\mid  & & m    &\in & \Heaps       &\isdef& \Addrs \rightarrow \Words \\
    & & & &\exec\mid \entry\mid \rwx\mid \rwlx &                                             & \Phi &\in & \ExecConfs   &\isdef& \Regs \times \Heaps \\
    \gl&\in&\Globals & ::= &\glob \mid \local &                                              & \ms  &\in & \MemSegments &::= & \Addrs \parfun \Words \\
    &&\Confs &::=& \multicolumn{6}{l}{\ExecConfs + \{\failed \} + \{\halted\} \times \Heaps}\\
     && \Caps  &\isdef & \multicolumn{6}{l}{\left\{
                         ((\perm,\gl),\start,\addrend,\addr)\mid
                         \start,\addr\in\Addrs,\addrend \in
                         \Addrs\cup \{\infty\}\right\}}
  \end{array}
\]
  \begin{equation*}
  \begin{array}{rcl}
    \rv    &\in& \ints + \RegName \\
    i      &::=& 
                 \jmp{\lv} \mid 
                 \jnz{\lv}{\rv} \mid
                 \move{\lv}{\rv} \mid 
                 \load{\lv}{\hv} \mid 
                 \store{\hv}{\rv} \mid
                 \lt{\lv}{\rv}{\rv} \mid
                 \plus{\lv}{\rv}{\rv} \mid 
                 \minus{\lv}{\rv}{\rv} \mid \\
          &   &  \lea{\lv}{\rv} \mid 
                 \restricttwo{\lv}{\rv} \mid
                 \subseg{\lv}{\rv}{\rv} \mid  
                 \isptr{\lv}{\rv} \mid 
                 \getl{\lv}{\lv} \mid
                 \getp{\lv}{\lv} \mid
                 \getb{\lv}{\lv} \mid\\
          &   &  \gete{\lv}{\lv} \mid 
                 \geta{\lv}{\lv} \mid 
                 \fail \mid
                 \halt 
  \end{array}
\end{equation*}
  \caption{The syntax of our capability machine assembly language.}
  \label{fig:syntax}
\end{figure}

\begin{figure}[htb]
  \centering
  \begin{equation*}
    \Phi  \rightarrow
    \begin{cases}
      \sem{\decode(n)}(\Phi) & \arraycolsep=0pt
      \begin{array}[t]{l}
        \text{if $\memreg(\pcreg) = \stdcap$ and $\start \leq \addr \leq \addrend$}\\ 
        \text{\;\;and $\perm \in \{ \exec,\rwx, \rwlx \}$ and $\memheap(\addr) = n$ }
      \end{array}
\\
      \failed                                 & \text{otherwise}
    \end{cases}
  \end{equation*}
  \begin{align*}
    \stdUpdatePc{\Phi} &=
                         \begin{cases}
                           \updateReg{\pcreg}{\var{newPc}} & \arraycolsep=0pt
                           \begin{array}[t]{l}
                             \text{if $\memreg(\pcreg) = \stdcap$}\\
                             \text{\;\;and $\var{newPc} = ((\perm,\gl),\start,\addrend,\addr + 1)$}
                           \end{array}
\\
                             \failed & \text{otherwise}
                         \end{cases}
  \end{align*}
  \begin{tabular}{|c|p{3.4cm}|>{\raggedright\arraybackslash}p{7.3cm}|}
    \hline
    $i$&$\sem{i}(\Phi)$&Conditions\\
    \hline 
    $\fail$&$\failed$&\\
    \hline
    $\halt$&$(\halted,\memheap)$&\\
    \hline
    $\move{\refreg{r_1}}{r_2}$& $\stdUpdatePc{}(\updateReg{r_1}{w})$&$r_2 \in \Regs \Rightarrow w = \memreg(r_2)$ and $r_2 \in \ints \Rightarrow w = r_2$\\
    \hline
    $\load{\refreg{r_1}}{\refheap{r_2}}$&$\stdUpdatePc{\updateReg{r_1}{w}}$&$\memreg(r_2) = \stdcap{}$ and  $w = \memheap(\addr)$ and $\start \leq \addr \leq \addrend$ and $\perm \in \{ \rwx, \rwlx, \exec, \readwrite, \readwritel, \readonly \}$ \\
    \hline
    $\restricttwo{\refreg{r_1}}{r_2}$&$\stdUpdatePc{}(\updateReg{r_1}{w})$  & $\memreg(r_2) = \stdcap$ and $(\perm',g') = \decodePermPair{\memreg(r_2)}$
                                                                              and $(\perm',g') \sqsubseteq (\perm,g)$ and  $w =((\perm',g'),\start,\addrend,\addr)$\\
    \hline
    \itoplas{$\subseg{\refreg{r_1}}{r_2}{r_3}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{w}}}$} &  \itoplas{$\memreg(r_1) = \stdcap$ and for $i \in \{2,3\}$ $n_i = \memreg(\rv_i)$ and $n_2 \in \nats$ and $\start \leq n_2$ and $n_3 \leq \addrend$ where either $n_3 \in \nats$ or ($n_3=\inftyend$ and $\addrend = \infty$) and $\perm \neq \entry$ and $w = ((\perm,\gl),n_1,n_2,\addr)$}\\
    \hline
    $\jmp{\lv}$&$\updateReg{\pcreg}{\var{newPc}}$& if $\memreg(r) = ((\entry,\gl),\start,\addrend,\addr)$, then $\var{newPc} = ((\exec,\gl),\start,\addrend,\addr)$ otherwise $\var{newPc} = \memreg(r)$\\
    \hline
    \itoplas{$\lea{\refreg{r_1}}{r_2}$} & \itoplas{$\stdUpdatePc{\updateReg{r_1}{\var{c}}}$} &\itoplas{$\memreg(r_1) = \stdcap$ and $n = \memreg(r_2)$ and $n \in \ints $ and $\perm \neq \entry$ and $\var{c} = ((\perm,\gl),\start,\addrend,\addr + n)$} \\
    \hline
    $\geta{\refreg{r_1}}{\refreg{r_2}}$ & $\stdUpdatePc{\updateReg{r_1}{\addr}}$ &
                                                $\memreg(r_2) = ((\_,\_),\_,\_,\addr)$\\
    \hline
    $\store{\refheap{r_1}}{\refreg{r_2}}$&$\stdUpdatePc{\updateHeap{\addr}{\var{w}}}$&$\memreg(r_1) = \stdcap$ and $\perm \in \{ \rwx, \rwlx, \readwrite, \readwritel\}$  and $\start \leq \addr \leq \addrend$ and $\var{w} = \memreg(r_2)$
                                                                                       and if $\var{w} = ((\_,\local),\_,\_,\_)$, then $\perm \in \{\rwlx,\readwritel \}$\\
    \hline
    \multicolumn{3}{|c|}{$\cdots$}\\
    \hline
    \_&$\failed$&otherwise\\
    \hline
  \end{tabular}
  \caption{An excerpt from the operational semantics.}
  \label{fig:op-sem}
\end{figure}

@@

\dominique{Follow order of figure in discussion?}
The $\load{}{}$ instruction loads the contents of the requested memory location
into a register, but only if the capability has appropriate authority (i.e.\
read permission and an appropriate range). The $\restricttwo{}{}$ instruction updates a
capability's permissions and global/local tag in the register file, but only if the new permissions are weaker than the original. 
%\itoplassug{Reviewer E, popl, would like $\sqsubseteq$ and $\decodePermPair{}$ explained. If we explain all of these details, this section becaomes a bit long and cluncky. (see tex for suggestion)}
%\itoplassug{As the permissions and global/local tag are not words on the machine, we decode the word in register $r_2$ as a pair of permission and locality which is done with the function $\decodePermPair{}$.}
% Reviewer E, popl: Figure 3: how is decodePermPair defined, and what is the relation between (perm', g') and (perm, g)?
It also never turns local capabilities into global ones.
The $\subseg{}{}{}$ instruction reduces the range of authority of a capability.
In order to represent the unbounded address space, we use $-42$ to represent infinity.

The $\jmp{}$ instruction updates the program counter to a requested location,
but it is complicated by the presence of \emph{enter capabilities}, modeled
after the M-Machine's~\citep{Carter:1994:HSF:195473.195579}. Enter capabilities
cannot be used to read, write or execute and their address and range cannot be
modified. They can only be used to jump to, but when that happens, their
permission changes to $\exec$. They can be used to represent a kind of closures:
an opaque package containing a piece of code together with local encapsulated
state. Such a package can be built as an enter capability $c =
((\entry,\gl),\start,\addrend,\addr)$ where the range $[\start,\addr-1]$
contains local state (data or capabilities) and $[\addr,\addrend]$ contains
instructions. The package is opaque to an adversary holding $c$ but when $c$ is
jumped to, the instructions can start executing and have access to the local
data through the updated version of $c$, now in the $\pcreg$-register. 
\lau{As we talked about what is described here, is not how we make closures in
the TR, so we should consider whether we should refrain from using that word
here. }

The instruction $\lea{}{}$ manipulates the current address of non enter-capabilities.
It is fine for a capability to have a current address outside its range of authority as long as it is not used with an instruction that requires a specific capability.
The instruction $\geta{}{}$ queries the current address of a capability and stores it in a register.

Finally, the $\store{}{}$ instruction updates the memory to the requested value
if the capability has write authority for the requested location. However, the
instruction is complicated by the presence of \emph{local capabilities}, modeled
after the ones in the CHERI processor~\citep{Watson2015Cheri}. Basically, local
capabilities are special in that they can only be kept in registers, i.e.\ they
cannot be stored to memory. This means that local capabilities can be
\emph{temporarily} given to an adversary, for the duration of an invocation: if
we make sure to clear the capability from the register file after control is
passed back to us, they will not have been able to store the capability.
However, there is one exception to the rule above: local capabilities can be
stored to memory for which we have a capability with write-local authority
(i.e.\ permission $\rwl$ or $\rwlx$). This is intended to accommodate a stack,
where register contents can be stored, including local capabilities. As long as
all capabilities with write-local authority are themselves local and the stack
is cleared after control is passed back by the adversary, we will see that this
does not break the intended behavior of local capabilities.

@@

\textbf{Program start-up} We assume that the language runtime initializes the
memory as follows: a contiguous array of memory is reserved for the stack, for
which we receive a stack pointer in the register $r_\stk$. We stress that
the stack is not built-in, but merely an abstraction we put on this piece of the
memory. The stack pointer is local and has $\rwlx$ permission. Note that this
means that we will be placing and executing instructions on the stack.
Crucially, the stack is the only part of memory for which the runtime (including
malloc, loading, linking) will ever provide $\rwlx$ or $\rwl$ capabilities.
Additionally, our examples typically also assume some memory to store
instructions or static data. Another part of memory (called the heap) is
initially governed by malloc and at program start-up, no other code has
capabilities for this memory. Malloc hands out $\rwx$ capabilities for allocated
regions as requested (no $\rwlx$ or $\rwl$ permissions). For simplicity, we
assume that memory allocated through malloc cannot be freed.

@@

In one of the registers, we also provide a return pointer, which must be a local
capability. If it were global, the adversary would be able to store away the
return pointer in a global data structure (i.e.\ there exists a global capability for it), and jump to it later, in
circumstances where this should not be possible. For example, they could store
the return pointer, legally jump to it a first time, wait to be invoked again
and then jump to the old return pointer a second time, instead of the new return
pointer received for the second invocation. Similarly, they could store the
return pointer, invoke a function in our code, wait for us to invoke them again
and then jump to the old return pointer rather than the new one, received for the
second invocation.
% \lau{If we need space, then maybe one of the examples is sufficient.}
% \lau{Both of these are good to have their as the two last examples of \sectionname~\ref{sec:examples} use them.}
%
% \lau{In the examples, we refer to the local return pointer as
%   ``protected return pointer''. We should either introduce that here
%   or use a different formulation in the examples. }
% \lau{It seemed did not seem to add anything to adopt the terminlogy - we
%   often highlight that the return pointer is local which in most cases
%   is so important that we do not want to leave it out.}
By making the return pointer local, we prevent such attacks.
The adversary can only store local capabilities with a write-local capability, and the only piece of memory governed by a write-local capability is the stack.
Since the stack pointer itself is also local, it can also only be stored on the stack.
There is no way for the adversary to recover either of these local capabilities because we clear the part of the stack that they had access to before we pass control back to them.

Note that storing stack pointers for use during future invocations would also be
dangerous in itself, i.e. not just because it can be used to store return
pointers. Imagine the adversary stores their stack pointer (\figurename~\ref{fig:illu-adv-store-sp}), invokes trusted code
that uses part of the stack to store private data (\figurename~\ref{fig:illu-adv-call}) and then invokes the adversary
again with a stack pointer restricted to exclude the part containing the private
data (\figurename~\ref{fig:illu-adv-called}). If the adversary had a way of keeping hold of their old stack pointer, it
could access the private data stored there by the trusted code and break
local-state encapsulation.  
% \lau{Do we say anything about arguments to the callback? Specifically, it should not be okay to uncritically pass on argument we recieved from an adversary. Can we even to any extend allow local arguments that we pass on from an unknown source?}
% Domi: this is now discussed in the discussion section, no need to discuss it here?

@@

Clearly, neither the old program counter or the old stack pointer should be accessible by the adversary. In
other words, the return pointer provided to the adversary must be a capability
that they can jump to but not read from, i.e.\ an enter capability. To make this
work, we construct the activation record as depicted in
\figurename~\ref{fig:activ-rec-struct}. The $\entry$ return pointer has authority
over the entire activation record (containing the previous return and stack
pointer), and its current address points to a number of restore instructions in
the record, so that upon invocation, these instructions are executed and can
load the old stack pointer and program counter back into the register file. As
the return pointer is an enter pointer, the adversary cannot get hold of the
activation record's contents, but after invocation, its permission is updated to
$\exec$, so the contents become available to the restore instructions.

@@

\paragraph{Local capabilities}
The last thing we provide informal intuition about is how our results handle local capabilities and their special treatment by the hardware.
Normally, when we invoke an untrusted piece of code and provide it with certain global capabilities, it may have stored those capabilities in memory and we will only be able to reinvoke the code if we can guarantee that those values are still valid.
Formally, worlds represent the invariants that global capabilities' safety relies on and the reinvocation is only safe in future worlds, where the invariants are respected.

@@

      \tikzset{capa/.style={decoration={markings,mark=at position 0 with %
            {\fill[black] circle(1.5pt);}}}}
      \draw[capa] (8.5,6.5) edge[-latex,postaction={decorate},out=180,in=0] node[above] {\tiny{$(p,\gl)$}} (5,7.25);

      \node[anchor=west] at (7.75,2) {$c_\var{next}$};
      \draw[capa] (7.75,2) edge[-latex,postaction={decorate},out=180,in=0] node[right] {\tiny{$(p,\gl)$}}  (5.4,5.26);
    \end{tikzpicture}
    \caption{Visualization of ``$(\reg,\ms)$ is looking at $[i_0\dots i_n]$
      followed by $c_{\var{next}}$''.}
  \label{fig:def-intuition-looking-at}
  \end{subfigure}
  \begin{subfigure}{.12\linewidth}
$\phantom{te}$
  \end{subfigure}
  \begin{subfigure}{0.43\linewidth}
    \begin{tikzpicture}[scale=.5, every node={scale=.5}]
    % recurrent parts
      \scope
      \clip (-.1,-.1) rectangle (4.6,11.1);
      \fill[fill=white] (0,0) rectangle (4.5,11);
      \fill[color=gray!20] (0,9.5) rectangle (4.5,11);
      \fill[color=gray!20] (0,0) rectangle (4.5,2);
      \draw (0,0) -- (0,11);
      \draw (4.5,0) -- (4.5,11);
